### 2024-08-19 article/article.tex
allT

\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{U1}
\author{Rodriguez González José Adrián }
\date{August 2024}

\begin{document}

\maketitle

\section{Abstract}
In data analysis we need to understand that every analysis problems requires its own time before to create a predictive model. This is because we need to verify if our model would it be useful.
We have two problems to practice that it has been learned about data analysis. As first checkout it'll be a simple dataset that only have two features \(x and y\). So at the first sight seems to be quick but before to create a predictive model, it'll be required to take a look into the data as first step. And also, we got another case about predicte the price of cars.
\section{Introduction}
\section{Related works}
\section{Methods and materials}
\section{Data analysis}
\section{Discussion}
\section{Conclussions and future work}
\section{references}

\end{document}